---
title: "HW 4"
author: "Ashley Cooper, Chloe Swick, and Eleanor Thomas"
date: "June 2, 2023"
output: 
  html_document:
    code_folding: hide
    theme: yeti
    collapsed: yes
    smooth_scroll: yes
---

```{r setup, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(patchwork)
library(cowplot)
library(equatiomatic)

warming <- read.csv(here("data", "warming.csv"))
damages <- read.csv(here("data", "damages.csv"))

### set one-time pulse in carbon emissions
pulse <- 35000000000
```

# Question 1
```{r fig.cap="Figure 1"}
### add a column to damages with warming squared
damages$warming_sq <- damages$warming^2

### linear regression model using the new warming_sq and the original warming 
quad_damages_lm <- lm(damages ~ warming + warming_sq -1, data = damages)

### damages function

quad_damages_fun <- function(x){
  quad_damages_lm$coefficients[1]*x+quad_damages_lm$coefficients[2]*x^2
}

### extract equation from damages model 
extract_eq(model = quad_damages_lm, use_coefs = TRUE, coef_digits = 2)

### plot 
ggplot(data = damages, aes(x = warming, y = damages)) +
  geom_point() +
  theme_minimal() +
  labs(x = "Warming (ºC)", y = "Damages ($)", title = "Damages vs. Global Mean Temperature Change") +
  stat_function(fun = quad_damages_fun, color = "lightblue2")

```


# Question 2
```{r fig.cap="Figure 2"}
### predict damages in each year under the baseline climate trajectory and under the pulse scenario
### add columns for baseline damages and pulse damages using damages function
### then take the difference and difference/per ton CO2 from the pulse
predicted_damages <- warming %>%
  mutate(damages_baseline = quad_damages_fun(warming_baseline)) %>%
  mutate(damages_pulse = quad_damages_fun(warming_pulse)) %>%
  mutate(damages_diff = damages_pulse - damages_baseline) %>%
  mutate(damages_diff_co2 = damages_diff/pulse)

### plot the baseline damages
plot_damages_no_pulse <- ggplot(data = predicted_damages, aes(x = year, y = damages_baseline)) +
  geom_line(color = "gold") +
  theme_minimal() +
  labs(x = "Year", y = "Baseline Damages ($)")

### plot the damages under the pulse scenario
plot_damages_pulse <- ggplot(data = predicted_damages, aes(x = year, y = damages_pulse)) +
  geom_line(color = "green4") +
  theme_minimal() +
  labs(x = "Year", y = "Damages + Pulse ($)")

### plot the difference between baseline and pulse scenarios
plot_damages_diff <- ggplot(data = predicted_damages, aes(x = year, y = damages_diff)) +
  geom_line(color = "pink") +
  theme_minimal() +
  labs(x = "Year", y = "Difference in Damages + Pulse ($)")

### plots the difference in damages over the pulse tons CO2
plot_damages_diff_co2 <- ggplot(data = predicted_damages, aes(x = year, y = damages_diff_co2)) +
  geom_line(color = "lightblue") +
  theme_minimal() +
  labs(x = "Year", y = "Difference in Damages + Pulse ($/ton CO2)")

(plot_damages_no_pulse+plot_damages_pulse)/(plot_damages_diff+plot_damages_diff_co2)
```

# Question 3
```{r fig.cap = "Figure 3"}
### set the baseline year
baseline_year <- 2021

### calculate the present value of damages for each year with discount rates 1%-8%
scc <- predicted_damages %>%
  select(year, damages_diff_co2) %>%
  mutate(pv_1 = damages_diff_co2*((1.01)^(-(year-baseline_year)))) %>%
  mutate(pv_2 = damages_diff_co2*((1.02)^(-(year-baseline_year)))) %>%
  mutate(pv_3 = damages_diff_co2*((1.03)^(-(year-baseline_year)))) %>%
  mutate(pv_4 = damages_diff_co2*((1.04)^(-(year-baseline_year)))) %>%
  mutate(pv_5 = damages_diff_co2*((1.05)^(-(year-baseline_year)))) %>%
  mutate(pv_6 = damages_diff_co2*((1.06)^(-(year-baseline_year)))) %>%
  mutate(pv_7 = damages_diff_co2*((1.07)^(-(year-baseline_year)))) %>%
  mutate(pv_8 = damages_diff_co2*((1.08)^(-(year-baseline_year))))
  
### sum the damages from 2023-2100 to get scc for each discount rate
scc_calc <- c(sum(scc$pv_1), sum(scc$pv_2),sum(scc$pv_3),sum(scc$pv_4),sum(scc$pv_5), sum(scc$pv_6),sum(scc$pv_7),sum(scc$pv_8))

### create a dataframe with the summed scc calculations and discount rates
disc_rates <- c(1:8)
scc_df <- data.frame(disc_rates,
                     scc_calc)

### plot the scc by discount rate, add in ramsey scc from question 4 (150.43)
### (put this here or question 4 or?)
ggplot(data = scc_df, aes(x = disc_rates, y = scc_calc)) +
  geom_point(color = "black") +
  theme_minimal() +
  labs(x = "Discount Rate (%)", y = "Social Cost of Carbon ($/ton of CO2)") +
  geom_vline(xintercept=2.1, color = "turquoise") +
  geom_label(aes(x=2.1, label="Ramsey Rule: SCC = $150.43", y=90), colour="turquoise", angle=90) +
  geom_point(aes(x = 2.1, y = 150.43), color = "turquoise", size = 2.5)
```


# Question 4

```{r}
### define given parameters
rho <- 0.001
n <- 2
g <- 0.01

### calculate ramsey discount rate
ramsey_dr <- (rho+n*g)*100

### calcualte predicted damages with ramsey discount rate 
predicted_damages_ramsey <- predicted_damages %>%
  select(year, damages_diff_co2) %>%
  mutate(pv_ramsey = damages_diff_co2*((1+(ramsey_dr/100))^(-(year-baseline_year))))

### sum damages to get ramsey scc, see figure 3
scc_ramsey <- sum(predicted_damages_ramsey$pv_ramsey)
```

# Question 5 

Now suppose there are two possible climate policies that can be pursued. Policy A is business as usual emissions, without any substantial climate change mitigation. Policy B is to take immediate and strong action on climate change. Use these facts:

If you undertake Policy A there are two possible outcomes. Either warming will occur as in the “baseline” (i.e. “no-pulse”) dataset above (this happens with probability 0.5) or warming each year will be 1.5 times that in the “baseline” dataset (with probability 0.5).

Under Policy B, warming will continue until 2050 as in the “baseline” dataset, and then will stabilize at 1.29 degrees and stay that way forever.

Society is risk neutral

Use a discount rate of 2%

### What is the expected present value of damages up to 2100 under Policy A?

```{r}
### Create columns a1, a2, b; calculate the discounted rate

# predicted_damages <- warming %>%
#   mutate(damages_baseline = quad_damages_fun(warming_baseline)) %>%
#   mutate(damages_pulse = quad_damages_fun(warming_pulse)) %>%
#   mutate(damages_diff = damages_pulse - damages_baseline) %>%
#   mutate(damages_diff_co2 = damages_diff/pulse)

policy <- warming %>% 
  select(-warming_pulse) %>% 
  mutate(policya1 = warming_baseline) %>% 
  mutate(policya2 = warming_baseline*1.5) %>% 
  mutate(policyb = warming_baseline) %>% 
  mutate(damagesa1 = predicted_damages$damages_baseline) %>% 
  mutate(damagesa2 = quad_damages_fun(policya2))

policy$policyb[policy$policyb > 1.29] <- 1.29

policy <- policy %>% 
  mutate(damagesb = quad_damages_fun(policyb))


```

#### What is the expected present value of damages up to 2100 under Policy B?
```{r}

```

#### Suppose undertaking Policy A costs zero and undertaking Policy B costs X. How large could X be for it to still make economic sense to pursue Policy B instead of Policy A?

```{r}

```

#### Qualitatively, how would your answer change if society were risk averse?

